Celll Cycle Analysis using Whitefield's microarray data (03/19/2019)

Read in all Whitefield's reference data, gene list and select only dataset 3 which contains most samples
```{r}
data <- read.table("./dataPlusScores_all5.remove.quote.txt",sep="\t",header=T,row.names = 1,fill=T)
cellcycle.genes <- read.table("./CellCycleGeneList_1134.txt",header=T,row.names = 1,sep="\t")
data.cc <- data[rownames(cellcycle.genes),]
data.cc3 <- data.cc[,c(52:99)]
data.cc3 <- data.cc3[rowSums(is.na(data.cc3))==0,]
```

Read in Gene ID conversion table--  MatchMiner is used to convert IMAGE clone ID into symbol ID. One IMAGE clone ID may correspond to multiple gene symbols and vice versa.
```{r}
id <- read.table("./genesymbol.all.list",header=F,sep="\t")
imageGene <- t(as.matrix(table(id)))
```

Rank genes according to atan2 value (This ranking doesn't really affect the result)
```{r}
library(ComplexHeatmap)
data.cc3.atan2 <- data.cc3[order(data.cc[rownames(data.cc3),]$Atan2),]
phases <- cellcycle.genes[rownames(data.cc3.atan2),]$PHASE
ha_column = HeatmapAnnotation(df = data.frame(type1 = phases), 
                              col = list(type1 = c("G1/S" =  "red", "G2" = "orange", "S phase"="yellow","*S phase"="yellow","G2/M"="green","M/G1"="blue")))

ht1 = Heatmap(t(as.matrix(data.cc3.atan2)),top_annotation = ha_column,show_row_names = F,show_column_names = F,cluster_rows = FALSE, cluster_columns = FALSE)

draw(ht1)
```

A new matrix ranked by the peaktime of genes for one single cycle.
```{r fig.width=10, fig.height=8}
cc <- 0:46
pt <- cc%%15.4
data.cc3.sub <- data.cc3[,2:48]
oatan2 <- data.cc[rownames(data.cc3),]$Atan2
ratan2 <- oatan2 + (oatan2 < 0)*2*pi
data.cc3.sub.rank <- data.cc3.sub[order(ratan2),order(pt)]
colnames(data.cc3.sub.rank) <- c(1:47)
phases <- cellcycle.genes[rownames(data.cc3.sub.rank),]$PHASE
ha_column = HeatmapAnnotation(df = data.frame(type1 = phases), 
                              col = list(type1 = c("G1/S" =  "red", "G2" = "orange", "S phase"="yellow","*S phase"="yellow","G2/M"="green","M/G1"="blue")))

ht1 = Heatmap(t(as.matrix(data.cc3.sub.rank)),top_annotation = ha_column,show_row_names = T,show_column_names = F,cluster_rows = FALSE, cluster_columns = FALSE)

draw(ht1)
```

Define normalization function
```{r}
normal <- function(countsdata){
  countsdata <- countsdata[which(apply(countsdata,1,sum)!=0),]
  countsdata <- countsdata[,which(apply(countsdata,2,sum)!=0)]
  data.norm <- countsdata
  cellsum <- colSums(countsdata)
  med=median(cellsum)
  for(i in c(1:dim(countsdata)[2])){
    data.norm[,i] <- (countsdata[,i]/cellsum[i])*med
  }
  data.log <- log10(data.norm+1)
  data.log
}
```

Process single cell data and infer cell cycle stages
```{r}
#Read in data

#Extract gene symbol in both of our data and conversion table and do conversion
genes <- intersect(rownames(spg.cca@raw.data),colnames(imageGene))
ig.matrix <- imageGene[,genes]
ig.matrix <- ig.matrix[rowSums(ig.matrix)!=0,]

spg.cells.ord <- spg.cca@cell.names[order(spg.cca@meta.data$res.0.6.ord)]
spg.cells.ord <- spg.cells.ord[spg.cca@meta.data[spg.cells.ord,]$res.0.6.ord %in% c(1:6)]
gc.matrix <- spg.cca@raw.data[genes,spg.cells.ord]
gc.matrix <- gc.matrix[,colSums(gc.matrix)!=0]
gc.matrix <- gc.matrix[rowSums(gc.matrix)!=0,]
spg.matrix <- as.matrix(ig.matrix) %*% as.matrix(gc.matrix)
spg.matrix <- spg.matrix[rowSums(spg.matrix)!=0,]
spg.matrix <- spg.matrix[,colSums(spg.matrix)!=0]

#format ref data
imageids <- intersect(rownames(spg.matrix),rownames(data.cc3.sub.rank))
ref <- data.cc3.sub.rank[imageids,]

#Normalization
spg.matrix <- normal(spg.matrix[imageids,])
for(i in c(1:dim(spg.matrix)[1])){
  spg.matrix[i,]=(spg.matrix[i,]-mean(spg.matrix[i,]))/sd(spg.matrix[i,])
}

#compute correlation matrix
corr <- cor(ref,spg.matrix,method = "spearman")

#Draw heatmap
ha_column = HeatmapAnnotation(df = data.frame(Clusters = as.vector(spg.cca@meta.data[spg.cells.ord,]$res.0.6.ord)), 
                              col = list(cellClusters = c("1" =  "#477900", "2"="#9378ff","3"="#01ce74","4" =  "#ff2b85", "5"="#b0d266","6"="#733697")))
ht1 = Heatmap(as.matrix(corr),top_annotation = ha_column, show_row_names = T,show_column_names = F,cluster_rows = FALSE, cluster_columns = FALSE)

draw(ht1)
```

```{r fig.width=10, fig.height=8}

#Infer stage by highest correlation
bupt <- rownames(corr)[max.col(t(corr))]
#cell.cycle <- data.frame(cell=colnames(data123),cycle=bupt)

#cells ordered by inferred stages
spg.cells.ord.inf <- spg.cells.ord[order(bupt)]

#Draw heatmap
ha_column = HeatmapAnnotation(df = data.frame(Clusters = as.vector(spg.cca@meta.data[spg.cells.ord.inf,]$res.0.6.ord)), 
                              col = list(Clusters = c("1" =  "#477900", "2"="#9378ff","3"="#01ce74","4" =  "#ff2b85", "5"="#b0d266","6"="#733697")))
ht2 = Heatmap(as.matrix(corr[,spg.cells.ord.inf]),top_annotation = ha_column, show_row_names = T,show_column_names = F,cluster_rows = FALSE, cluster_columns = FALSE)

draw(ht2)
```

show in the t-SNE plot
```{r fig.width=6.5, fig.height=4.5}
spg.cca@meta.data$cellcycle <- 0
spg.cca@meta.data[spg.cells.ord,]$cellcycle <- as.numeric(bupt)
TSNEPlot(object = spg.cca, do.return = TRUE, pt.size = 1, group.by = "cellcycle", cells.use = spg.cca@cell.names[spg.cca@meta.data$res.0.6.ord %in% c(1:6)])
```

